---
title: "01-visualizar-datos-presentacion"
author: "Renan Escalante-Chong"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: ioslides_presentation
---

```{r setup1, include=FALSE}
library(knitr)
library(ggplot2)
```

# Visualización de datos

## Introducción

Este tutorial te mostrará como visualizar datos usando la librería de R ggplot2.
Este paquete fue creado for Hadley Wickman.
La mayoría de los ejemplos de este tutorial están basados en el libro [R for data science](http://r4ds.had.co.nz)

## Data frames

Un data frame es una estructura de datos en R
Un data frame es una tabla donde las columnas (columns) son variables y las filas (rows) son instancias, casos.

```{r, include=FALSE}
x<-c("Juan","Mildred","Sofía")
y<-c(22,18,21)
fac<-c("UNAM","UV","preparatoria")
estado<-c("Tabasco","Campeche","Puebla")
d <- data.frame(estudiantes = x, edad = y, fac = fac,estado=estado)
```

Nota: Este tipo de datos son datos limpios **tidy data** donde las columnas son siempre variables y las filas son instancias.

```{r, echo=F}
kable(d)
```

## Prerequisites

To access the datasets, help pages, and functions that we will use in this chapter, load ggplot2 using the `library()` function. We'll also load tibble, which you'll learn about later. It improves the default printing of datasets.

```{r setup}
library(ggplot2)
```

If you run this code and get the error message "there is no package called ‘ggplot2’", you'll need to first install it, then run `library()` once again.

```{r eval = FALSE}
install.packages("ggplot2")
library(ggplot2)
```

## A graphing template

Let's use our first graph to answer a question: Do cars with big engines use more fuel than cars with small engines? 

You can test your answer with the `mpg` dataset in ggplot2, or `ggplot2::mpg`:
  
```{r, echo=FALSE}
mpg2<-as.data.frame(mpg[1:5,])
kable(mpg2)
```

```{r, include=FALSE}
#The dataset contains observations collected by the EPA on 38 models of car. Among the variables in `mpg` are:
#  
#  1. `displ`, a car's engine size, in litres.
#
#1. `hwy`, a car's fuel efficiency on the highway, in miles per gallon (mpg). 
#A car with a low fuel efficiency consumes more fuel than a car with a high 
#fuel efficiency when they travel the same distance. 
#
#To learn more about `mpg`, open its help page by running `?mpg`.
#
#To plot `mpg`, open an R session and run the code below. The code plots the `mpg` data by putting `displ` on the x-axis and `hwy` on the y-axis:
``` 

## Gráficos 

Muchas veces podemos usar números para tratar de entender un set de datos usando representaciones gráficas

Los autos con motores grandes consumen más gasolina?
  
## A graphing template (2)

Esta es una forma de describir un gráfico

```{r, fig.height = 4 , fig.width= 7}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```


## Gramática de gráficos

Aquí vamos a aprender la gramática de como construir un gráfico

1. Set de datos (data)
2. Especificar que variables y que instancias del set de datos queremos (mapping)
3. Luego especificamos que tipo de plot queremos (geom_point)
4. Por último agregamos propiedades del gráfico 

```{r eval=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```


## Exercises

1.  Run `ggplot(data = mpg)` what do you see?

1.  What does the `drv` variable describe?  Read the help for `?mpg` to find
    out.
     
1.  Make a scatterplot of `hwy` vs `cyl`.

1.  What happens if you make a scatterplot of `class` vs `drv`. Why is
    the plot not useful?
    
## Agregar "aesthetics" podemos agregar propiedades al gráfico

```{r, fig.height = 3 , fig.width= 7}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, 
                           color = class))
```

## Gráfico usando tamaño (mal ejemplo)

```{r, fig.height = 3 , fig.width= 7}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, 
                           size = class))
```

## Gráfico usando formas
```{r, fig.height = 3 , fig.width= 7, eval=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```

## Gráfico usando formas
```{r, fig.height = 3 , fig.width= 7, echo=FALSE, warning=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```

## Selección condicional (color)

```{r, fig.height = 3 , fig.width= 7}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = displ<5))
```

## Obtener ayuda

Escriban ?geom_point

```{r}
datos<-read.table("efectos_medicina_ALS.tsv", header = FALSE)
datos_ALS<-read.table("datos_ALS.tsv", header = FALSE)
```




